Vertical shear instability in astrophysical disks: vertically global linear calculations with thermal relaxation and applications to the Solar Nebula

### 2015-03-12 numerical.tex
updating results

%, so that we may consistently consider vertical domains up to
%$z\sim r$.

We solve the linear problem by expanding the
perturbations in Chebyshev polynomials $T_l$ up to $l=512$
and discretizing the equations on a grid with
$z\in[-\zmax,\zmax]$. At the vertical boundaries we impose a free
surface, 
\begin{align}
  \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0 \quad \text{at } z=\pm\zmax,
\end{align}
where $\bm{\xi}$ is the Lagrangian displacement, the meridional
components of which are given by $\xi_{x,z} = \ii\delta
v_{x,z}/\sigma$. Calculations with a rigid boundary ($\delta v_z=0$)
yield similar results. 
% This boundary
% condition is convenient for numerical implementation since it conforms
% to a standard eigenvalue problem.

@@

% mention mode selection -> focus on fundamental corrugation ->
% simulations eventually dominated by them 
% choice of kx is O(10), since nelson observe kxH \sim 30
%not possible to survey all modes/kx/disk parameters
%focus on effect of thermal relaxation

\subsection{VSI with rapid thermal relaxation}\label{vertiso_pertiso} 
% disk model is a nearly-vertically
% isothermal disk with $\Gamma=1.011$, $p=-1.5$, $q=-1$,
% $\epsilon=0.1$. 
We first calculate the VSI in a neutrally
stratified disk ($N_z^2\equiv0$) with rapid thermal relaxation by
setting $\gamma=\Gamma=1.011$ and $\beta=10^{-3}$. Other disk
parameters are $(p,q,\epsilon)=(-1.5,-1,0.1)$ and the vertical domain
is $\zmax = 5H$. The disk is nearly vertically isothermal with
$c_s(\zmax)\simeq 0.94 c_s(0)$. Fig. \ref{omega_z} shows that the 
equilibrium rotation decreases slightly away  from the disk mid-plane.

This setup is similar to that adopted
in \cite{mcnally14}, who calculated the VSI in vertically isothermal
disks subject to isothermal perturbations ($\Gamma=\gamma=1$).  This
allows us to test our numerical code. We also compare results with our 
analytical discussion of isothermal perturbations in Appendix
\ref{iso_discuss}.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  0cm]{figures/omega2} 
  \caption{Equilibrium rotation profile $\Omega(z)$,
    normalized by its mid-plane value, for a disk with $\Gamma=1.011$
    and $(p,q,\epsilon)=(-1.5,-1,0.1)$. 
    \label{omega_z} 
  }
\end{figure}

%behaviour as func of kx

Fig. \ref{iso_eigen_kx} compares the numerically-obtained eigenvalues
for the fundamental mode and that calculated from 
Eq. \ref{sig2_iso} (with $L=1$). There is good agreement for 
$\khat\gtrsim10$. The match worsens for smaller $\khat$ because the 
the low-frequency approximation used to obtain Eq. \ref{sig2_iso} 
is no longer valid, as $|\omega|$ is not small compared to $\Omega_k$
for $\khat\lesssim 10$.

@@

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/compare_eigen_imag_iso} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  1cm]{figures/compare_eigen_real_iso}
  \caption{Growth rate (top) and real frequency (bottom) of the
    fundamental VSI mode in a disk with $\gamma=\Gamma=1.011$,
    $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
    $\beta=10^{-3}$. The solid (dotted) line is obtained numerically
    (analytically).  
    \label{iso_eigen_kx} 
  }
\end{figure}

Fig. \ref{lowfreq_eigenfunc} compares the fundamental VSI eigenfunction
$W$ for $\khat=1$ and $\khat=20\pi$. The latter value, corresponding
to $k_x= 200\pi/r$, was used by \cite{mcnally14}. We find $W$ is
similar for all $\khat\gtrsim10$ and is proportional to $z$, as
expected for the fundamental mode according to explicit solutions
developed in Appendix \ref{iso_poly}. For completeness we show the
corresponding eigenfunctions in vertical velocity in
Fig. \ref{lowfreq_eigenfuncvz}. The fundamental mode
consists of zero nodes in $\delta v_z$.

%For $\khat\lesssim 10$, e.g. $\khat=1$ in 
%Fig. \ref{lowfreq_eigenfunc}, the perturbation magnitude near the
%boundaries begin to dominate. For even smaller $\khat\lesssim 1$ (not
%shown) we find the perturbation is almost entirely concentrated at the
%boundaries. However, our radially-local approximation is not valid for 
%such small wavenumber perturbations.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/eigenvector_iso_kx1} 
  \includegraphics[width=\linewidth]{figures/eigenvector_iso_kx60}
  \caption{Eigenfunction $W$ of the fundamental VSI for perturbation
    wavenumber $\khat=1$ (top) and $\khat=20\pi$ (bottom). The real 
    (imaginary) parts of $W$ are plotted as solid (dotted) lines. % The
    % eigenfunction has been normalized such that $\imag(W)=0$ at the
    % vertical boundaries and $\mathrm{max}|W|=1$. 
    \label{lowfreq_eigenfunc}
  }
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/eigenvectorvz_iso_kx1} 
  \includegraphics[width=\linewidth]{figures/eigenvectorvz_iso_kx60}
  \caption{Same as Fig. \ref{lowfreq_eigenfunc} but for the vertical
    velocity eigenfunctions.
    \label{lowfreq_eigenfuncvz}
  }
\end{figure}

In Fig. \ref{lowfreq_eigen} we plot the eigenvalues 
$\sigma = \omega + \ii\nu$ found for $\khat=20\pi$. This plot is
similar to Fig. 3 in \cite{mcnally14}. The fundamental mode has the
smallest $|\sigma|$, for which the eigenvalue calculated from
Eq. \ref{sig2_iso} with $L=1$ and the numerically-obtained value are   
\begin{align*}
  (\omega,\nu) = (-0.3053,0.2606)\epsilon\Omega_k &\quad \text{(from
    Eq. \ref{sig2_iso})},\\
  (\omega,\nu) = (-0.3010,0.2575) \epsilon \Omega_k &\quad \text{(numerical)}.
\end{align*}
This agreement is surprisingly good given the number of approximations
used to obtain Eq. \ref{sig2_iso}, which also imposes a different
boundary condition to that in the numerical calculation. These values
are also close to that found by \cite{mcnally14}. The eigenfrequency
is not sensitive to the vertical boundary condition because the
expression for $\sigma$, given by Eq. \ref{integral_relation1},
involve integrals with $\rho$ as the   weight function, which rapidly
decays away from the midplane.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/eigenvalues_iso}
  \caption{Eigenvalues of perturbations with wavenumber $\khat=20\pi$
    in a neutrally-stratified, nearly-vertically isothermal disk
    ($\gamma=\Gamma=1.011$) with $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
    $\beta=10^{-3}$. \label{lowfreq_eigen} 
  }
\end{figure}

\subsection{Effect of thermal relaxation}
We now consider stably stratified disks ($\gamma > \Gamma$) with
thermal relaxation. Here, we use a slightly  
different disk model with $\epsilon=0.05$ and $\gamma=1.4$, as 
adopted in simulations performed by \cite{nelson13}. Other parameters 
are unchanged from the previous section.

@@

% Fig. \ref{relax_growth_num} shows the growth rates of the fundamental  
% mode as a function of $\khat$. This plot is largely consistent with  
% Fig. \ref{relax_disp_fig} except for $\khat\lesssim 1$
% where the low-frequency approximation fails. (In any case, the local model not valid
% for $\khat\ll1$.) For $\khat\gtrsim 10$ introducing thermal
% relaxation rapidly stabilizes the fundamental mode.

% Fig. \ref{relax_growth_num} also confirms our analytical discussion that
% introducing a small but finite cooling time stabilizes the fundamental
% VSI, but there is a preferable wavenumber that maximizes this effect (\S\ref{relax_pert}).
% This occurs at $\khat= O(10)$ for the current disk model.

% \begin{figure}
%    \includegraphics[width=\linewidth,clip=true,trim=0cm 0.0cm 0cm
%    0cm]{figures/compare_eigen_imag_bloop} 
%   % \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
%   % 1cm]{figures/compare_eigen_imag_bloop2} 
%   \caption{Growth rates of the fundamental VSI mode in 
%     nearly-vertically isothermal disks ($\Gamma=1.011$) with
%     $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ under a
%     range of thermal relaxation timescales   
%     $\beta$. The eigenfrequencies are calculated by numerically 
%     solving the full linear eigenvalue problem. \label{relax_growth_num}}   
% \end{figure}

%approximately the perturbation wavenumber observed in the fiducial
%numerical simulation of \cite{nelson13}.

We visualize the effect of thermal relaxation in 
Fig. \ref{relax_eigenW_num} by comparing the vertical velocity
perturbation $\delta v_z$ for
$\beta=0.01$ and $\beta=0.1$. The latter value is close to
$\beta_\mathrm{crit}=0.125$. 
% the
% critical value beyond which the fundamental VSI is stabilized
% (\S\ref{iso_vsi_beta_crit}). 
Increasing $\beta$ 
restricts  the region in which $\delta v_z\sim $ constant closer to the midplane, with
oscillatory behavior emerging at the boundaries, where buoyancy first
becomes important. Finite thermal relaxation increases $|\delta v_z|$ near the
vertical boundaries relative to that about the midplane. However, 
perturbations near the disk surface may not be energetically important
in practice because the disk atmosphere contains little mass.

% We also find this trend for fixed $\beta$ but
% increasing 
% $\khat$.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/eigenvectorvz_beta0d01} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  0cm]{figures/eigenvectorvz_beta0d1} 
  \caption{Numerically-calculated fundamental VSI
    mode in a nearly vertically isothermal disk ($\Gamma=1.011$) with
    $(p,q,\epsilon)=(-1.5,-1,0.05)$, evolved with $\gamma=1.4$ and a dimensionless 
    thermal relaxation timescale $\beta = 0.01$
    (top) and $\beta=0.1$ (bottom). The
    real (imaginary) part of the vertical velocity perturbation
    $\delta v_z$ is plotted as the solid (dotted)
    line. % The eigenfunction has been normalized such that
    % $\imag(\delta v_z)=0$ at the vertical boundaries, and
    % $\mathrm{max}|\delta v_z|=1$.
    \label{relax_eigenW_num}}  
\end{figure}

\subsection{Critical thermal relaxation timescale}\label{bcrit_num_test}
% may need to explain why choose k=10
% bcrit independent of k for k>>1 
% for some parameter regimes (e.g. large q, large epsilon) 
% behavior at larger kx and finite bcool not understood. 
We test the dependence of the critical thermal relaxation timescale 
$\beta_\mathrm{crit}$ for the fundamental VSI on disk parameters
(Eq. \ref{iso_vsi_cond}).  We use the previous case study as a
reference point and vary $q\in[-0.2,-1.2]$, 
$\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$
separately. Here, we use a wavenumber $\khat=10$, for which we
find growth rates reach zero for the range of disk parameters
considered, so that a numerical $\beta_\mathrm{crit}$ can be defined precisely.

>>>>

We solve the linear problem by expanding the
perturbations in Chebyshev polynomials $T_l$ up to $l=512$
and discretizing the equations on a grid with
$z\in[-\zmax,\zmax]$. Our standard boundary conditions at the vertical
boundaries is a free surface, 
\begin{align}
  \Delta P \equiv \delta P + \bm{\xi}\cdot\nabla P= 0 \quad \text{at } z=\pm\zmax,
\end{align}
where $\bm{\xi}$ is the Lagrangian displacement, the meridional
components of which are given by $\xi_{x,z} = \ii\delta
v_{x,z}/\sigma$.

@@

Unless otherwise stated, our fiducial disk model is: a nearly 
vertically isothermal disk with $\Gamma=1.011$, stable
stratification $\gamma=1.4$, $(p,q,\epsilon)=(-1.5,-1,0.05)$, and 
vertical domain size $\zmax=5H$. In Fig. \ref{omega_z} we plot the
corresponding equilibrium rotation profile, which shows that $\Omega$
decreases slightly away from the mid-plane.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  0cm]{figures/omega2} 
  \caption{Equilibrium rotation profile $\Omega(z)$,
    normalized by its mid-plane value, for a disk with $\Gamma=1.011$
    and $(p,q,\epsilon)=(-1.5,-1,0.05)$. 
    \label{omega_z} 
  }
\end{figure}

\subsection{VSI with rapid thermal relaxation}\label{vertiso_pertiso} 
We first calculate the VSI in a disk with rapid thermal relaxation by
setting $\beta=10^{-3}$. (Smaller values yield the same results.) This
is similar to strictly isothermal perturbations in a vertically
isothermal disk considered in \cite{nelson13} and \cite{mcnally14}. % We
% compare results with our analytical discussion of isothermal 
% perturbations in Appendix \ref{iso_discuss}.

Fig. \ref{iso_eigen_kx} compares the numerically obtained eigenvalues 
for the fundamental mode and that calculated analytically from our
analysis of isothermal perturbations in Appendix \ref{iso_discuss}
(Eq. \ref{simple_growth} with $L=1$). We find good agreement in the
wave frequency $\omega$ at all $\khat$ and in growth rates $\nu$ for
$\khat\gtrsim 10$. Growth rates are $O(\epsilon\Omega_k)$, 
consistent with our discussion in Appendix \ref{max_growth1}, but
are over-estimated by our analysis. The match in $\nu$ worsens for
decreasing $\khat$ due to the neglect of global radial structure in
the  analysis. Nevertheless, given the number of approximations made
to calculate growth rates analytically, this agreement is
satisfactory, especially at large $\khat$.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/compare_eigen_imag_iso} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  1cm]{figures/compare_eigen_real_iso}
  \caption{Growth rate (top) and real frequency (bottom) of the
    fundamental VSI mode in a disk with $(\gamma,
    \Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
    $\beta=10^{-3}$. The solid (dotted) line is obtained numerically
    (analytically).  
    \label{iso_eigen_kx} 
  }
\end{figure}

@@

Fig. \ref{lowfreq_eigenfunc} plots the eigenfunctions $W$ and $\delta 
v_z$ for the fundamental mode with $\khat=10$ (i.e. the mode with
maximum growth rate in Fig. \ref{iso_eigen_kx}). We find $W\propto z$
throughout most of the disk, corresponding to a constant vertical
velocity perturbation, as expected for the fundamental mode according
to explicit solutions developed in Appendix \ref{iso_explicit}. We
show a meridional visualization of this mode in
Fig. \ref{lowfreq_eigenfunc_2d}. Although there is a constant vertical
velocity throughout the fluid column, most of the momentum is
contained within $\sim 2H$ of the midplane due to the density
stratification.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/eigenvectorW_iso} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  1cm]{figures/eigenvectorvz_iso}
  \caption{Pressure perturbation $W$ (top) and vertical velocity
    perturbation $\delta v_z$ of the fundamental VSI for a
    perturbation  wavenumber $\khat=10$  The real 
    (imaginary) parts of $W$ and $\delta v_z$ are plotted as solid
    (dotted) lines. 
    \label{lowfreq_eigenfunc}
  }
\end{figure}

\begin{figure}
  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 3.1cm
  0cm]{figures/result2d_vel}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
  0cm]{figures/result2d_mom} 
  \caption{Visualization of the fundamental VSI mode in
    Fig. \ref{lowfreq_eigenfunc}. The color scale corresponds to the
    real pressure perturbation $W$ scaled by its maximum value.
    Arrows in the left panel correspond to the meridional velocity
    field $(\delta v_x,\delta v_z)$, and arrows in the right panel
    correspond to to the meridional momentum $\rho(\delta v_x,\delta
    v_z)$. 
    \label{lowfreq_eigenfunc_2d}
  }
\end{figure}

\subsubsection{Case study: $\khat=10$}
In Fig. \ref{compare_modes_iso_kx10} we plot eigenvalues of unstable
modes found for our standard setup with $\khat=10$. This case is shown
by red crosses. We also plot several other calculations with different
vertical boundary conditions: free surface at $\zmax = 3H$ (black
dots), free surface at $\zmax=7H$ (green triangles) and a rigid
boundary at $\zmax=5H$ (blue cross).

The fundamental mode, marked by `f', corresponds to the eigenvalue
with smallest magnitude $|\sigma|$. We see that the fundamental mode,
while not the most unstable in this case, is more robust against
vertical boundary conditions. Higher order modes with larger growth
rates (i.e. the branch where $\nu$ increases with $|\omega|$) require
larger vertical domains.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx10.ps}
  \caption{Unstable modes for a disk with 
    $(\gamma,\Gamma)=(1.4,1.011)$, $(p,q,\epsilon)=(-1.5,-1,0.1)$ and
    $\beta=10^{-3}$. The perturbation wavenumber is $\khat=10$. 
    The standard setup with a free surface ($\Delta P=0$) and 
    vertical domain size $\zmax=5H$ is shown by the red cross. Other
    cases correspond to different boundary condtions applied at
    different heights. The fundamental mode is marked by `f'.   
    \label{compare_modes_iso_kx10} 
  }
\end{figure}

\subsubsection{Surface modes: $\khat=30$} 
Our standard setup with $\khat=10$ in 
Fig. \ref{compare_modes_iso_kx10} (red crosses) differs to that from
\cite{nelson13} and \cite{mcnally14} in the lack of `surface modes' in
our case. These modes lie parallel to the vertical axis in mode
diagrams such as Fig. \ref{compare_modes_iso_kx10}. We find them for
sufficiently large $\zmax$ and/or $\khat$. % For example, the two
% largest growth rates found for the case with $\zmax=7H$ in
% Fig. \ref{compare_modes_iso_kx10} are surface modes.

To illustrate this, we plot the mode diagram for $\khat=30$ in  
Fig. \ref{compare_modes_iso_kx30} with different boundary 
conditions. Surface modes can be seen with
$\zmax\geq 5H$, examples of which are marked by `s'. These modes have
been discussed in the above studies. They typically have the largest
growth rates but require sufficiently large vertical domain size and their eigenvalues
vary with $\zmax$.

Due to their dependence on boundary conditions, we
do not consider surface modes further in detail. In fact, we will see
in the next section that surface modes are more effectively stabilized
by finite thermal relaxation than the fundamental mode. %  Thus, surface
% modes may not be the most relevant in realistic disks.

\begin{figure}
  \includegraphics[width=\linewidth]{figures/compare_modes_iso_kx30_tags.ps}
  \caption{Same as Fig. \ref{compare_modes_iso_kx10} but for $\khat=30$
    \label{compare_modes_iso_kx30}. Examples of surface modes are
    marked with `s'. These are found for sufficiently large vertical
    domain sizes. 
  }
\end{figure}

\subsection{Effect of thermal relaxation}
We now consider the effect of thermal relaxation by increasing
$\beta$. Note that for the standard disk model the critical thermal 
timescale $\beta_\mathrm{crit} = 0.125$.

Fig. \ref{compare_modes_cool_kx10} show unstable modes with $\khat=10$
and $\beta\in[0.01,0.1]$. In addition to our standard setup with 
$\zmax=5H$, we also plot results for $\zmax=3H$ and $\zmax=7H$. In all
three cases the number of unstable modes decrease with increasing
$\beta$. The fundamental mode is not the most unstable for 
$\beta=0.01$ as there is a large number of higher order modes, but
these depend on boundary conditions.

However, the fundamental mode is dominant for $\beta \geq 
0.03$. Higher order modes, with larger growth rates than the
fundamental mode, are more effectively stabilized with increasing 
$\beta$. For $\beta=0.1$, which is close to $\beta_\mathrm{crit}$, we
only find the fundamental mode.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/compare_modes_cool_kx10_z3.ps} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/compare_modes_cool_kx10_z5.ps}
  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx10_z7.ps}
  \caption{Unstable modes with $\khat=10$ and thermal
    relaxation timescales $\beta=0.01$ (black dots), $\beta=0.03$ (red
    crosses), $\beta=0.05$ (green triangles) and $\beta=0.1$ (blue
    crosses); for different vertical domain sizes $\zmax=3H$ (top),
    $\zmax=5H$ (middle, standard setup) and $\zmax=7H$ (bottom). The
    fundamental mode is marked with `f'. 
    \label{compare_modes_cool_kx10} 
  }
\end{figure}

Fig. \ref{compare_modes_cool_kx30} show unstable modes with
$\khat=30$. In this case surface modes appear for sufficiently large   
$\zmax$. However, they are effectively stabilized with increasing 
$\beta$. For $\zmax=3H$, the fundamental mode becomes dominant for
$\beta\geq 0.03$, as before. For $\zmax=5H$ and $\zmax=7H$, we find
the appearence of modes slightly more unstable than the fundamental
mode even when $\beta=0.1$ (the leftmost eigenvalues in the lower two
panels). However, these modes are clearly associated with the vertical
domain size as they are absent for $\zmax=3H$. % Thus, for $\zmax=5H$ and 
% $\zmax=7H$  the fundamental mode is likely dominant for 
% $\beta\geq0.05$

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/compare_modes_cool_kx30_z3.ps} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/compare_modes_cool_kx30_z5.ps}
  \includegraphics[width=\linewidth]{figures/compare_modes_cool_kx30_z7.ps}
  \caption{Same as Fig. \ref{compare_modes_cool_kx10} but for
    $\khat=30$. Examples of surface modes are marked with `s', and 
    appear for sufficiently large $\zmax$.
    \label{compare_modes_cool_kx30} 
  }
\end{figure}

Fig. \ref{compare_modes_cool_kx10} and \ref{compare_modes_cool_kx30} 
show that the fundamental mode is not sensitive to boundary
conditions. This suggests that tracking fundamental mode would be a
more robust way of assessing stability as a function of $\beta$,
rather than searching for the most 
unstable mode, because the latter may be associated with boundary
conditions. In fact, the fundamental mode bcomes dominant 
(or close to it) with increasing $\beta$.

\begin{figure}
  \includegraphics[width=\linewidth,clip=true,trim=0cm 1.75cm 0cm
  0cm]{figures/eigenvectorW_beta0d1} 
  \includegraphics[width=\linewidth,clip=true,trim=0cm 0cm 0cm
  1cm]{figures/eigenvectorvz_beta0d1}
  \caption{Same as Fig. \ref{lowfreq_eigenfunc_cool} but with a
    thermal relaxation timescale $\beta=0.1$. 
    \label{lowfreq_eigenfunc_cool}
  }
\end{figure}

\begin{figure}
  \includegraphics[scale=0.345,clip=true,trim=0cm 0cm 3.1cm
  0cm]{figures/result2d_vel_cool}\includegraphics[scale=0.345,clip=true,trim=1.9cm 0cm 0cm
  0cm]{figures/result2d_mom_cool} 
  \caption{Same as  Fig. \ref{lowfreq_eigenfunc_2d} but with a thermal
    relaxation timescale $\beta=0.1$. 
    \label{lowfreq_eigenfunc_2d_cool}
  }
\end{figure}

\subsection{Critical thermal relaxation
  timescale}\label{bcrit_num_test}

@@

% may need to explain why choose k=10
% bcrit independent of k for k>>1 
% for some parameter regimes (e.g. large q, large epsilon) 
% behavior at larger kx and finite bcool not understood. 
We test the dependence of the critical thermal relaxation timescale 
$\beta_\mathrm{crit}$ for the fundamental VSI on disk parameters
(Eq. \ref{iso_vsi_cond}).  We use the previous case study as a
reference point and vary $q\in[-0.2,-1.2]$, 
$\gamma\in[1.2,2.0]$, and $\epsilon\in[0.02,0.1]$
separately. Here, we use a wavenumber $\khat=10$, for which we
find growth rates reach zero for the range of disk parameters
considered, so that a numerical $\beta_\mathrm{crit}$ can be defined precisely.
